Parallelization of Push-based System for Molecular Simulation Data Analysis with GPU

### 2016-03-01 paper/paperSource.tex
automatic triggered changes

\emph{\textbf{One-body queries kernel}}.
Commonly used CUDA array reduction\cite{gpureduction} makes the implementation of all the one-body queries obvious. Figure \ref{fg:reduction} nicely represents general idea of the data mapping, where each cell of the vector can be associated with a thread of execution in the kernel. Having an array of atom data structures, we reduce the array accumulating all the needed data from each atom's attributes. We don't expect much performance improvement from coalesced memory access, since the data structure for atom can eliminate any gain of coalesced memory access, though for specific types of queries it could be possible to aggregate certain attributes into merged arrays and take an advantage of it. In this case, the solution should be simple in order to keep it easily extendable for other simple one-body queries for a scientist to program. We fetch all the atoms from GPU global memory, since we use references to atoms once. In this case, allocating shared memory and copying it over would complicate the implementation and possibly create additional overhead for one time reference usage. Finally, memory transportation and data consumption are implemented for user, thus, he only needs to extend the queries module which consequently can be turned into plugins model.

@@

\subsection{Results}
Results are nicely represented, in general, on Figures \ref{fg:chart_1000}, \ref{fg:chart_10000} \ref{fg:chart_max}. You can instantly notice that speedup for 1000 atoms is not very impressive, especially for $10$ frames, and the reason for that is the following. There is a small overhead we have to consider before computing part. In order to let GPU process the data, we have to transfer the data to Global Memory, and upon finish we have to load the results back to RAM. In this particular case, for small workload and frames, it is even slower than the sequential processing. Afterwards, the bigger workload the more convincing speedup gets, unfortunately, it doesn't get more than $4$ times for 1000 atoms. Fortunately, scientists usually work with much larger data sets, thus, this doesn't really represent general speedup for average use case.

@@

@@

\section{Conclusions and Future Work}\label{sc:conclusion}
The objective of our work is to design and implement improved data
analysis system that can be used in the field of molecular
simulation system's analysis. In this paper, we introduce the idea
for such system. We build our system on a push-based type design,
where data from data arrays is being pushed onto available queries
in the system. These queries are being executed on the pushed data
and produce intermediate / final result that would be used as part
of the data analysis. We are able to achieve an improvement over
existing, pull-based type designs because of the I/O overhead such
designs introduce when dealing with large volumes of scientific
data. Also, our queries are able to be executed on the same stream
of data, making it suitable solution for streaming circumstances. We
designed a benchmark that can be used to test data analysis systems.
This benchmark comprises of three parts: 1)benchmark data,
2)benchmark queries, and 3) benchmark parameters. We use this
benchmark to compare our system to one of the most frequently used
MS analysis systems, Gromacs. The efficiency and speedup achieved by
our system is supported by extensive experiments and their results.
The results show that our push-based design achieves up to about
$1000$ times speedup in comparison to a pull-based design, i.e.,
Gromacs.

One direction of our future work will be to further improve our
push-based design. Through the extensive experiments we have learned
that our design can be improved when the atom selection clause
involves many conditions. This improvement may be in the direction
of improving the algorithmic part, but it can also be in the
direction of improving the data presentation/organization we have
used in the system.\\

>>>>

\emph{\textbf{One-body queries kernel}}.
Commonly used CUDA array reduction\cite{gpureduction} makes the implementation of all the one-body queries obvious. Figure \ref{fg:reduction} represents general idea of the data mapping, where each cell of the vector can be associated with a thread of execution in the kernel. Having an array of atom data structures, we reduce the array accumulating all the needed data from each atom's attributes. We don't expect much performance improvement from coalesced memory access, since the data structure for atom can eliminate any gain of coalesced memory access, though for specific types of queries it could be possible to aggregate certain attributes into merged arrays and take an advantage of it. In this case, the solution should be simple in order to keep it easily extendable for other simple one-body queries for a scientist to program. We fetch all the atoms from GPU global memory, since we use references to atoms once. In this case, allocating shared memory and copying it over would complicate the implementation and possibly create additional overhead for one time reference usage. Finally, memory transportation and data consumption are implemented for user, thus, he only needs to extend the queries module which consequently can be turned into plugins model.

@@

\subsection{Results}
Results are represented, in general, on Figures \ref{fg:chart_1000}, \ref{fg:chart_10000} \ref{fg:chart_max}. You can instantly notice that speedup for 1000 atoms is not very impressive, especially for $10$ frames, and the reason for that is nature of GPU and memory transfer. There is a small overhead we have to consider before computing part: in order to let GPU process the data, we have to transfer the data to Global Memory, and upon completion it is needed to load the results back to RAM to show it to the user. In this particular case, for small workload and frames, it is even slower than the sequential processing. Afterwards, the bigger workload the more convincing speedup gets, unfortunately, it doesn't get more than $4$ times for 1000 atoms. Fortunately, scientists usually work with much larger data sets, thus, this doesn't really represent general speedup for average use case.

@@

Fortunately, scientific data is much bigger than above examples. You can see average speedup for bigger number of atoms on Figure \ref{fg:chart_max}. The speedup gets upto 70 times which could save huge amount of time for scientists. These results are not the best, as it has been already mentioned earlier, since we have not modified general structure of memory being transferred to Global Memory not taking advantage of coalesced memory access to make it easily extendable.

@@

It is noticable that growth of the speedup is not stable and with certain number of atoms it drops dramatically. It is because we've used static approach of computing grid dimensions. Besides it, in the future, it can be expanded into a dynamic version that will automatically adjust the grid dimensions choosing least overhead that it has to go with.

\section{Conclusions and Future Work}\label{sc:conclusion}
The ultimate goal of this work was to get an improved version of Push-based system for molecular simulation data analysis. As you could see from result on charts (Figure \ref{fg:chart_1000}, \ref{fg:chart_max}), the proposed improved version of the system could save a lot of time for scientists which is up to 60-70 times less of time. The results have been gathered with a single snapshot of the framework tracked in control version system which can be improved with means of data converter module and dynamic grid dimensions computation module. In the future, this system has all potentials to be expanded into a distributed system, where frames could be aggregated by nodes and processed in parallel.\\
